Sampling Strategies in Decisions from Experience

### 2022-04-08 manuscript.Rmd
Review of Structure: Moved CPT review into prediction section

abstract: "In decisions from experience, systematic choices between prospects are based on the assessment of differences in latent properties. Yet, several properties can be used to distinguish among the prospects. To this end, this paper presents a sampling and accumulation model that allows sampling  strategies—i.e., the probability of sampling outcomes from different prospects in direct succession (switching probability)—to alter which differences in the latent properties evidence is accumulated for. Specifically, the proposed model sequentially samples outcomes from two prospects and accumulates the results obtained from a sequence of ordinal comparisons between sample means, with switching probabilities determining the size of the samples underlying each comparison and thresholds determining how much comparisons a prospect must win in order to be chosen. A computational implementation of the model is used to simulate the respective information-processing chains upfront choices between a safe and a two-outcome prospect, while varying switching probabilities and thresholds. The analysis of the simulation data shows that, depending on whether the accumulation process unfolds over mean comparisons based on small or large samples, evidence is accumulated for differences either in the probability to produce a higher outcome most of the time or in the expected value. The respective choice patterns reflect an as-if underweighting of rare outcomes or expected value maximization, both of which become more robust with increasing thresholds. The analysis moreover shows how the distinct choice patterns resulting from differences in the information-processing chains translate to the parameters of cumulative prospect theory’s value and weighting function."

@@

Now, let a *prospect* be a choice alternative that possesses only two kinds of fundamental properties: the possible outcomes of the alternative (e.g., gains or losses of some amount) and the probabilities with which these outcomes occur following the choice of the alternative.
Importantly, the information that the inputs provide about these properties can come in different forms.
Specifically, in *decisions from description* (DFD), the inputs take the form of explicit and complete descriptions of all outcomes and probabilities.
Yet, in *decisions from experience* [DFE, @hertwigDecisionsExperienceEffect2004], these properties are latent and therefore not known with certainty, but they must be inferred from the relative frequencies with which the outcomes occurred in the past, e.g., when the prospect was chosen in similar past decisions.
The inputs to DFE therefore take the form of sampled outcomes.
Although these distinct input forms may in principle carry the same information about a given set of prospects, behavioral decision research has so far produced a large body of papers indicating that---taking a quite general perspective---there are robust differences between the choices that are made in DFD and DFE, leading to the notion of the *description-experience gap* [@hertwigDescriptionexperienceGapRisky2009].
In one reading of the gap,^[<!--Open footnote-->See @hertwigConstructBehaviorGap2018, for a discussion on the interpretation of the gap.<!--Close footnote-->]
choice patterns in DFD deviate from the solutions of expected value (EV) maximization, as if small-probability outcomes are given more weight than would be warranted by their objective probabilities;
in turn, in DFE, choice patterns deviate from the solutions of EV maximization, as if the same small-probability outcomes are given less weight than would be objectively warranted [e.g., @barronSmallFeedbackbasedDecisions2003; @hertwigDecisionsExperienceEffect2004; @weberPredictingRiskSensitivity2004; @erevAnomaliesForecastsDescriptive2017; @regenwetterConstructBehaviorGap2017; see @wulffMetaanalyticReviewTwo2018, for a comprehensive meta-analytic review].
Hereafter, these choice patterns are referred to as the *as-if overweighting* and *as-if underweighting of rare outcomes*, respectively.

@@

Therefore, consider first the finding that *sampling strategies* were indicative for the final choice in DFE.
Specifically, in one paradigm used to study DFE, namely, the sampling paradigm [@weberPredictingRiskSensitivity2004], before making a final choice between prospects with latent properties, people can, as much as they want, sequentially sample the prospects' outcomes according to their objective probabilities.
Considering this paradigm and the choice between two prospects, @hillsInformationSearchDecisions2010 distinguished sampling strategies along the frequency of switches between the two prospects during the sampling phase,^[<!--Open footnote-->Note that @hillsInformationSearchDecisions2010 proposed two paradigmatic sampling strategies, where prospects are either switched just once (comprehensive sampling) or after each sampled outcome (piecewise sampling).
Nevertheless, in their analyses of empirical sampling data, they used the number of switches between prospects relative to the overall number of sampled outcomes from both prospects (switching frequency) as an approximate measure for the two proposed sampling strategies, and further acknowledged that "many [sampling] strategies will fall on the continuum between [the two paradigmatic strategies]" [@hillsInformationSearchDecisions2010, p. 1788].
Respectively, in the current paper sampling strategies are defined as switching probabilities.
<!--Close footnote-->] 
and proposed two different decision strategies that people may adopt.
That is, for a summary strategy, it is assumed that the prospect is chosen which produced the greater mean across all sampled outcomes.
In turn, for a round-wise strategy, it is assumed that over multiple rounds two sampled outcomes---i.e., one from each of the prospects---are compared and that the prospect is chosen which won the comparison in most rounds.
Then, @hillsInformationSearchDecisions2010 found that the choices of people who switched frequently between prospects during the sampling phase were better predicted by a round-wise strategy (rather than a summary strategy), and that this predictive pattern reversed for people who switched infrequently.

Now, one explanation for such a reversed predictive pattern would be if the differences in the sampling strategies determine which of the two decision strategies is used.
To this end, this paper presents a computational model that provides a link between *switching probabilities*---i.e., the probability of sampling outcomes from different prospects in direct succession during the sampling phase---and the decision strategies.
Specifically, the proposed model accumulates the outcomes obtained from the comparisons between sample means, with switching probabilities determining the size of the samples underlying each mean comparison and thresholds determining how much (more) of such comparisons a prospect must win in order to be chosen.
Accordingly, the model is such that the higher (lower) the switching probabilities and the higher (lower) the thresholds, the more the accumulation process resembles the round-wise (summary) strategy.

The current paper discusses and demonstrates in a computational analysis how such a simple accumulation model can explain an as-if underweighting of rare outcomes, even if a large number of sampled outcomes is fully accounted for.
More specifically, since the switching probabilities alter the size of the samples underlying each mean comparison, the model predicts the switching probabilities to affect whether a low-probability outcome contributes to the majority of required mean comparisons about as much as would be warranted by its objective probabilities, or whether it contributes rather less.
Respectively, if the small-probability outcome is smaller (larger) than the EV, high switching probabilities cause the mean to be an inflated (deflated) estimate of the EV in most of the comparisons, which can translate to an as-if underweighting of rare outcomes, irrespective of what the thresholds determine the required number of mean comparisons to be.
Yet, since the EV is not the only property which can be used to distinguish among the prospects, the proposed model shows how the mind may rely on different sampling strategies to change which differences in the latent properties of prospects evidence is accumulate for.
Specifically, for low switching probabilities, the accumulation process unfolds over mean comparisons based on large samples and is therefore more indicative for differences in the EV; in turn, for high switching probabilities, the accumulation process unfolds over mean comparisons based on small samples and is therefore more indicative for differences in the probability to produce a higher outcome than the respective other prospect most of the time.
In either case, then, the model implements a speed-accuracy trade-off of the form that the respective differences in the latent properties are more accurately assessed, the larger the thresholds are, i.e., the more evidence must be accumulated in favor of a prospect.

To substantiate the discussion, the model is used to simulate the accumulation process for different combinations of switching probabilities and thresholds and to predict choices between a two-outcome prospect and a safe prospect, respectively.
Considering choice rates and parameter estimates of cumulative prospect theory [CPT, @tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993], the computational analysis shows that the choices converge to the solutions of EV maximization for low switching probabilities, but take the form of an as-if underweighting of rare outcomes for high switching probabilities.
Importantly, the analysis also shows that the differences between switching probabilities become more pronounced with increasing thresholds.
In other words, the proposed model predicts an as-if underweighting of rare outcomes to occur when high switching probabilities cause the accumulation process to unfold over mean comparisons that are based on small samples.
Thus, the analysis substantiates the claim that not only the reliance on just a portion of a large number of sampled outcomes can explain an as-if underweighting of rare outcomes in DFE, but also the reliance on multiple or many small samples.
Moreover, the analysis shows that an as-if underweighting of rare outcomes cannot be equated with inaccurate inferences about latent properties---e.g., the EVs---that carry over to the subsequent choices.
Rather the opposite, an as-if underweighting of rare outcomes can also be caused by an accurate assessment of the differences among prospects in the latent probability to produce a higher outcome most of the time.
Before presenting the model and the computational analysis, the following section provides a formal definition of prospects and reviews how CPT in general and its weighting function in particular can be used to capture deviations from EV maximization.

# Prospects and the Description of Choices in CPT

Since there is an uncountable number of choices people confront in the natural world, behavioral decision research routinely abstracts from choices between particular choice alternatives, e.g., the choice between job offers, political parties to vote for, investment plans, and whatnot.
Rather, with the choice between at least two prospects, it studies a case which is counterfactual in that it omits the many particularities of each choice situation, but retains in the form of prospects the fundamental properties that almost all choice alternatives are assumed to possess.
These fundamental properties are considered the possible outcomes of a choice alternative and the probabilities with which these outcomes occur following the choice of the alternative.
Roughly speaking, then, all choice alternatives that can be fully described by their outcome-probability pairs are prospects.
As such, prospects can be understood as the distribution of a discrete random variable, for which the following section briefly provides an axiomatic basis.

## Formal Definition of Prospects

Following @kolmogorovFoundationsTheoryProbability1950 and adopting the symbolic notation from @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, let the tuple $(\Omega, \mathcal{F}, P)$ be a probability space.
Accordingly, $\Omega$ is a finite sample space with each element $\omega \in \Omega$ denoting a possible consequence of choosing the respective prospect.
$\mathcal{F}$ is the power set $\mathcal{P}(\Omega)$ with each element $A \in \mathcal{F}$ being a subset of $\Omega$, also denoted a random event.
$P$ is the probability measure
$$
P: \mathcal{F} \to [0,1]
\tag{1}
$$
assigning the elements of $\mathcal{F}$ a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.
Now, the random variable is a function
$$
\begin{aligned}
  X : (\Omega, \mathcal{F}) &\to (\Omega', \mathcal{F'})  \\
  \omega &\mapsto x
  \; ,
\end{aligned}
\tag{2}
$$
where $(\Omega', \mathcal{F'})$ is a measurable space with each element $x \in \Omega'$ denoting a possible outcome of a prospect and each element $A' \in \mathcal{F'}$ being a subset of $\Omega'$.
An outcome's probability $p_X(x)$ is then provided by the pushforward measure
$$
P' : \mathcal{F'} \to [0,1]
\tag{3}
$$
with 
$$
P'(A' \in \mathcal{F'}) := P(\{\omega: X(\omega) = x \in A'\})
\; .
\tag{4}
$$
Note that the measure in (3) and (4), respectively, is a random variable's distribution and therefore a prospect. 
Accordingly, given a choice between $1, ..., k$ prospects, the inputs from the environment or memory should provide information about the set
$$
\{P'^1, ... , P'^k\}
\; ,
\tag{5}
$$
where each element $P'$ can itself be considered a set of outcome-probability pairs $\{(x_i, p_X(x_i)\}_{i \in \mathbb{N}}$.

## CPT and the Concept of Probability Weighting

The choices people actually make between prospects are often described in terms of deviations from the principle of EV maximization, according to which the prospect with the largest EV
$$
EV = \sum_i^n x_i \times p_X(x_i)
\tag{6}
$$ 
should be chosen.
To *describe* how people's actual choices deviate from EV maximization [see @erevAnomaliesForecastsDescriptive2017, for a recent replication of classical demonstrations],^[<!--Open footnote-->@tverskyAdvancesProspectTheory1992 addressed the *fourfold pattern of risk attitudes*---i.e., choices between prospects indicate risk aversion for gains of high probability and losses of low probability, and risk seeking for gains of low probability and losses of high probability---and related effects, e.g., the *certainty effect* and the *reflection effect*, where the latter were already addressed by CPT's predecessor, prospect theory [@kahnemanProspectTheoryAnalysis1979]. Together, these empirical choice patterns indicate a violation of EV and expected utility (EU) maximization [see @bernoulliExpositionNewTheory1954\/1738].<!--Close footnote-->]
CPT and similar rank-dependent models [see @stottCumulativeProspectTheory2006, for an overview of models] fit choice data by assuming that people maximize the value
$$
V = \sum_i^n v(x_i) \times \pi_i
\; ,
\tag{7}
$$ 
with objective outcomes $x_i$ being transformed by a value function $v$, and *cumulative decision weights* $\pi_i$ being determined by the difference between transformed cumulative probabilities of the distribution $P'$.
More specifically, following @tverskyAdvancesProspectTheory1992, the objective outcomes are transformed by a value function
$$
\begin{aligned}
  v : \Omega' &\to \mathbb{R} \\
  x &\mapsto 
  \begin{cases}
     x_i^\alpha &\forall x_i \geq 0, \\
     -\lambda |x_i|^\alpha &\text{else}
     \; ,
  \end{cases}
\end{aligned}
\tag{8}
$$
with $\alpha \in [0,1]$ determining the degree of the function's concavity (convexity) over the positive (negative) outcome interval, and $\lambda > 1$ increasing the function's slope over the negative outcome interval only.
Each subjective value $v(x)$ is then multiplied (or: weighted) with a cumulative decision weight that takes the form
$$
\pi_i =
  \begin{cases}
     w^+(P(X \geq x_i)) - w^+(P(X > x_i)) \quad \forall x_i \geq 0, \\
     w^-(P(X \leq x_i)) - w^-(P(X < x_i)) \quad \text{else}
     \; , 
  \end{cases}
\tag{9}
$$
where $w$ is a monotonic increasing, nonlinear weighting function $w: [0,1] \to [0,1]$ satisfying $w^+(0) = w^-(0) = 0$ and $w^+(1) = w^-(1) = 1$.

Essentially, the model is such that the cumulative decision weights derived for the outcomes may be greater or smaller than their objective probabilities, causing the transformed outcomes---i.e., the subjective values---to be over- or underweighted *in* CPT, respectively.
Accordingly, for the description of choices, one may adopt the as-if weighting terminology from the introduction by stating that people choose as if they maximized the value in (7) and applied the weighting pattern that was estimated in CPT.
Importantly, the as-if prefix indicates that it cannot be concluded from the estimated weighting pattern that the mind does indeed perform any of the computations associated with the weighting of subjective values in CPT; rather, the mind processes the information about the choice alternatives in a way that the resulting choices translate to the estimated weighting pattern.
The as-if weighting terminology is therefore reminiscent of the distinction between different levels of explanation [cf. @marrVisionComputationalInvestigation1982].
That is, CPT may be considered a computational-level theory that specifies the computational problem the mind faces when being confronted with the choice between prospects, and how this problem may be solved by the abstract calculus of CPT.
However, as such, CPT cannot substitute algorithmic-level theories and process models that make claims about how exactly the mind transforms the inputs to the decision-making process into outputs which approximate the solutions of CPT [see @zilkerMeasuringModelingConstruction2020; see also @griffithsProbabilisticModelsCognition2010].
Respectively, this paper uses CPT as a mere data model to show how differences in the information-processing chain upfront choices in DFE---considered in the following section---may cause particular weighting patterns, among them the as-if underweighting of rare outcomes which has been found to be a robust behavioral phenomena in DFE.
To this end, the paper first continues to review the weighting patterns that can possibly estimated in CPT.

Now, several consequences for the weighting pattern that can be estimated in CPT follow from (9):

>>>>

abstract: "In decisions from experience, systematic choices between prospects are based on the assessment of differences in latent properties. Yet, several properties can be used to distinguish among the prospects. To this end, this paper presents a sampling and accumulation model that allows sampling  strategies—i.e., the probability of sampling outcomes from different prospects in direct succession (switching probability)—to alter which differences in the latent properties evidence is accumulated for. Specifically, the proposed model sequentially samples outcomes from two prospects and accumulates the results obtained from a sequence of ordinal comparisons between sample means, with switching probabilities determining the size of the samples underlying each comparison and thresholds determining how many comparisons a prospect must win in order to be chosen. A computational implementation of the model is used to simulate the respective information-processing chains upfront choices between a safe and a two-outcome prospect, while varying switching probabilities and thresholds. The analysis of the simulation data shows that, depending on whether the accumulation process unfolds over mean comparisons based on small or large samples, evidence is accumulated for differences either in the probability to produce a higher outcome most of the time or in the expected value. The respective choice patterns reflect an as-if underweighting of rare outcomes or expected value maximization, both of which become more robust with increasing thresholds. The analysis moreover shows how the distinct choice patterns resulting from differences in the information-processing chains translate to the parameters of cumulative prospect theory’s value and weighting function."

@@

Now, let a *prospect* be a choice alternative that possesses only two kinds of fundamental properties: the possible outcomes of the alternative (e.g., gains or losses of some amount) and the probabilities with which these outcomes occur following the choice of the alternative.
Importantly, the information that the inputs provide about these properties can come in different forms.
Specifically, in *decisions from description* (DFD), the inputs take the form of explicit and complete descriptions of all outcomes and probabilities.
Yet, in *decisions from experience* [DFE, @hertwigDecisionsExperienceEffect2004], these properties are latent and therefore not known with certainty, but they must be inferred from the relative frequencies with which the outcomes occurred in the past, e.g., when the prospect was chosen in similar past decisions.
The inputs to DFE therefore take the form of sampled outcomes.
Although these distinct input forms may in principle carry the same information about a given set of prospects, behavioral decision research has so far produced a large body of papers indicating that---taking a quite general perspective---there are robust differences between the choices that are made in DFD and DFE, leading to the notion of the *description-experience gap* [@hertwigDescriptionexperienceGapRisky2009].
In one reading of the gap,^[<!--Open footnote-->See @hertwigConstructbehaviorGapDescription2018, for a discussion on the interpretation of the gap.<!--Close footnote-->]
choice patterns in DFD deviate from the solutions of expected value (EV) maximization, as if small-probability outcomes are given more weight than would be warranted by their objective probabilities;
in turn, in DFE, choice patterns deviate from the solutions of EV maximization, as if the same small-probability outcomes are given less weight than would be objectively warranted [e.g., @barronSmallFeedbackbasedDecisions2003; @hertwigDecisionsExperienceEffect2004; @weberPredictingRiskSensitivity2004; @erevAnomaliesForecastsDescriptive2017; @regenwetterConstructbehaviorGapBehavioral2017; see @wulffMetaanalyticReviewTwo2018, for a comprehensive meta-analytic review].
Hereafter, these choice patterns are referred to as the *as-if overweighting* and *as-if underweighting of rare outcomes*, respectively.

@@

Therefore, first consider the finding that in DFE *sampling strategies* were indicative for the final choice.
That is, in one paradigm used to study DFE, namely, the sampling paradigm [@weberPredictingRiskSensitivity2004], before making a final choice between prospects with latent properties, people can, as much as they want, sequentially sample the prospects' outcomes according to their objective probabilities.
Considering this paradigm and the choice between two prospects, @hillsInformationSearchDecisions2010 distinguished sampling strategies along the frequency of switches between the two prospects during the sampling phase,^[<!--Open footnote-->Note that @hillsInformationSearchDecisions2010 proposed two paradigmatic sampling strategies, where prospects are either switched just once (comprehensive sampling) or after each sampled outcome (piecewise sampling).
Nevertheless, in their analyses of empirical sampling data, they used the number of switches between prospects relative to the overall number of sampled outcomes from both prospects (switching frequency) as an approximate measure for the two proposed sampling strategies, and further acknowledged that "many [sampling] strategies will fall on the continuum between [the two paradigmatic strategies]" [@hillsInformationSearchDecisions2010, p. 1788].
Respectively, in the current paper sampling strategies are defined as switching probabilities.
<!--Close footnote-->] 
and proposed two different decision strategies that people may adopt.
Specifically, for a summary strategy, it is assumed that the prospect is chosen which produced the greater mean across all sampled outcomes.
In turn, for a round-wise strategy, it is assumed that over multiple rounds two sampled outcomes---i.e., one from each of the prospects---are compared and that the prospect is chosen which won the comparison in most rounds.
Then, @hillsInformationSearchDecisions2010 found that the choices of people who switched frequently between prospects during the sampling phase were better predicted by a round-wise strategy (rather than a summary strategy), and that this predictive pattern reversed for people who switched infrequently.

Now, the current paper presents a sampling and accumulation model that provides an algorithmic link between sampling and decision strategies, and uses the model to discuss how the mind may rely on sampling strategies to alter which differences in the latent properties of prospects evidence is accumulated for.
More specifically, the proposed model assumes an accumulation process over a sequence of mean comparisons, with switching probabilities---i.e., the probability of sampling
outcomes from different prospects in direct succession---determining the size of the samples underlying each mean comparison and thresholds determining how many comparisons a prospect must win in order to be chosen.
The model is such that when switching probabilities are high, the accumulation process unfolds over mean comparisons based on small samples, thus resembling the round-wise strategy and accumulating evidence for differences in the probability to produce a higher outcome most of the time.
Notably, in this case, small-probability outcomes are predicted to contribute to the majority of mean comparisons less than would be objectively warranted, which can translate to an as-if underweighting of rare outcomes, irrespective of what the thresholds require the number of won comparisons to be.
<!--Accordingly, while increases in thresholds---and thus in the number of sampled outcomes distributed over multiple small samples, each of which is considered---allow to choose with increasing accuracy the prospect that is more likely to return a higher outcome, the resulting choices can produce an as-if underweighting of rare outcomes.-->
In turn, for low switching probabilities, the accumulation process unfolds over mean comparisons based on large samples, thus resembling the summary strategy---for low thresholds, respectively---and accumulating evidence for differences in the EV.

To substantiate the discussion, the model is used to simulate the sampling and accumulation process for choices between a safe and a two-outcome prospect, while systematically combining different switching probabilities and thresholds.
Considering choice rates and parameter estimates of cumulative prospect theory [@tverskyAdvancesProspectTheory1992; @wakkerAxiomatizationCumulativeProspect1993], the computational analysis shows that the choices converge to the solutions of EV maximization for low switching probabilities, but take the form of an as-if underweighting of rare outcomes for high switching probabilities.
Importantly, the analysis also shows that these distinct choice patterns become more robust with increasing thresholds.
Before presenting the model and the computational analysis, the following section provides a formal definition of prospects.

# Formal Definition of Prospects

Since there is an uncountable number of choices people confront in the natural world, behavioral decision research routinely abstracts from choices between particular choice alternatives, e.g., the choice between job offers, political parties to vote for, investment plans, and whatnot.
Rather, with the choice between at least two prospects, it studies a case which is counterfactual in that it omits the many particularities of each choice situation, but retains in the form of prospects the fundamental properties that almost all choice alternatives are assumed to possess.^[<!--Open footnote-->See @knightRiskUncertaintyProfit1921, Chapter 1, for a timeless discussion on the reasonableness of such abstractions.<!--Close footnote-->]
These fundamental properties are considered the possible outcomes of a choice alternative and the probabilities with which these outcomes occur following the choice of the alternative.
Roughly speaking, then, all choice alternatives that can be fully described by their outcome-probability pairs are prospects.
As such, prospects can be understood as the distribution of a discrete random variable.
More specifically, following @kolmogorovFoundationsTheoryProbability1950 and adopting the notation from @georgiiStochastikEinfuhrungWahrscheinlichkeitstheorie2015, let the tuple $(\Omega, \mathcal{F}, P)$ be a probability space.
Accordingly, $\Omega$ is a finite sample space with each element $\omega \in \Omega$ denoting a possible consequence of choosing the respective prospect.
$\mathcal{F}$ is the power set $\mathcal{P}(\Omega)$ with each element $A \in \mathcal{F}$ being a subset of $\Omega$, also denoted a random event.
$P$ is the probability measure
$$
P: \mathcal{F} \to [0,1]
\tag{1}
$$
assigning the elements of $\mathcal{F}$ a probability $0 < p(A) \leq 1$ with $P(\Omega) = 1$.
Now, the random variable is a function
$$
\begin{aligned}
  X : (\Omega, \mathcal{F}) &\to (\Omega', \mathcal{F'})  \\
  \omega &\mapsto x
  \; ,
\end{aligned}
\tag{2}
$$
where $(\Omega', \mathcal{F'})$ is a measurable space with each element $x \in \Omega'$ denoting a possible outcome of a prospect and each element $A' \in \mathcal{F'}$ being a subset of $\Omega'$.
An outcome's probability $p_X(x)$ is then provided by the pushforward measure
$$
P' : \mathcal{F'} \to [0,1]
\tag{3}
$$
with 
$$
P'(A' \in \mathcal{F'}) := P(\{\omega: X(\omega) = x \in A'\})
\; .
\tag{4}
$$
Note that the measure in (3) and (4), respectively, is a random variable's distribution and therefore a prospect. 
Accordingly, given a choice between $1, ..., k$ prospects, the inputs from the environment or memory should provide information about the set
$$
\{P'^1, ... , P'^k\}
\; ,
\tag{5}
$$
where each element $P'$ can itself be considered a finite set of outcome-probability pairs $\{(x_i, p_X(x_i)\}_{i \in \mathbb{N}}$.

# Sampling Strategies and the Accumulation of Evidence

Since it is only through their fundamental properties, or combinations thereof, that any distinction between the prospects can be made, systematic---as opposed to random---choices between prospects must be based on the assessment of differences in their properties.
However, only rarely in peoples' daily life, the inputs to the decision-making process take the form of explicit descriptions of all outcome-probability pairs.
In such cases, people would make DFD, since the information from which the mind learns about the properties of prospects are complete descriptions thereof.
Yet, rather often, people make DFE, where the mind can learn about the latent properties only by experiential sampling over time [@hertwigDescriptionexperienceGapRisky2009, p. 517].
Accordingly, for the choice between $1, ..., k$ prospects, the inputs to the decision-making process are generated by a set of prospect-specific stochastic processes
$$
\{\{X^1_t\}, ... , \{X^k_t\}\}
\; ,
\tag{6}
$$
where each $\{X_{t}\}_{t \in \mathbb N}$ is a collection of independent and identically distributed (i.i.d.) random variables.^[Note that it has been emphasized that in the natural world samples are "virtually never random" [@fiedlerBewareSamplesCognitiveecological2000, p. 660]. However, in the sampling paradigm, i.i.d. random variables are commonly assumed.]
The inputs to DFE are then the realizations of the random variables in (6).

The mind may use these realizations---i.e., the sampled outcomes---to first infer the latent properties and then proceed by selecting a property, or combinations thereof, to assess differences among the prospects and make a respective choice.
However, the mind may also approximate such a process by directly accumulating evidence about the differences in latent properties from the sampled outcomes.
To this end, sequential sampling models of decision making [e.g., @busemeyerDecisionFieldTheory1993] assume that the outcomes of prospects are sequentially sampled and integrated into dynamic decision variables and that a choice is made once a decision variable exceeds a threshold in favor for one of the prospects.
This class of process models thereby provides an algorithmic link between the actual sampling process and the choices in DFE, and has shown to explain some of the robust deviations from EV maximization [see @bhatiaSequentialSamplingParadoxes2014; @busemeyerDecisionFieldTheory1993], including the as-if underweighting of rare outcomes in the sampling paradigm [see @markantModelingChoiceSearch2015].
Moreover, these models implement speed-accuracy trade-offs, with the amount of evidence that must be sampled in favor for a prospect increasing with thresholds, leading to more (less) accurate but slower (faster) choices for high (low) thresholds.

Yet, there are several distinct properties that can be used to assess differences between the prospects, among them the EV, for which the summary decision strategy is indicative, and the probability to produce a higher outcome most of the time, for which the round-wise decision strategy is indicative [see @hillsInformationSearchDecisions2010].
Respectively, in the following, both decision strategies are implemented into a common sampling and accumulation model to demonstrate how the mind may rely on different sampling strategies to alter the degree to which evidence is accumulated for differences in either one of these properties.
More specifically, the proposed model accumulates the outcomes obtained from a sequence of comparisons between sample means, with the size of the samples underlying each mean comparison tending to be large (small) for low (high) switching probabilities.
Accordingly, whereas the decision variables of the model capture how many comparisons a prospect has won, the unfolding accumulation process more closely resembles the summary strategy for low switching probabilities and the round-wise strategy for high switching probabilities.

We consider the sampling paradigm and a choice between two prospects denoting the distributions $P'^X$ and $P'^Y$ of the discrete random variables $X$ and $Y$, respectively.
The model assumes that the sampling process starts at random with a stochastic process on one of the two random variables.
Specifically, it is assumed that an agent starts with equal probability to sample from one of the prospects, say $P'^X$, and generates a sequence of random variables $X_1, X_2,...$, where the mean over the sequence is computed or updated with each new outcome that is sampled.
The respective stochastic process $\{S^X_n\}_{n \in \mathbb{N}}$ is defined by
$$
S^X_n = \frac{1}{n} \sum_{t = 1}^n X_t, \quad n \in \mathbb{N}
\; ,
\tag{7}
$$
and terminates as soon as an outcome is sampled from the other prospect.
The switching probability $\psi \in (0,1]$, which is assumed to be adopted prior the start of the information-processing chain and is fixed throughout, controls the probability with which outcomes from different prospects are sampled in direct succession.
In other words, the switching probability is the probability with which the stochastic process on the current random variable stops and a new stochastic process on the other random variable starts with the subsequently sampled outcome.
To implement the accumulation process, the model assumes that each time after a new stochastic process was started and terminated on both random variables, their respective values are compared, with the prospect underlying the stochastic process with the greater final value receiving a round win.
Accordingly, for the prospect $P'^X$, this stage is modeled as the mapping
$$
\begin{aligned}
  Z^X : \mathbb{R} &\to \mathbb{N} \\
  S^X_n - S^Y_n &\mapsto 
  \begin{cases}
     1 &\text{if} \quad  S^X_n > S^Y_n, \\
     0 &\text{else}
     \; ,
  \end{cases}
\end{aligned}
\tag{8}
$$
where $Z^X$ is a random variable that takes only the two values $0$ and $1$, denoting a lost or won mean comparison, respectively. 
The model assumes that a round consisting of a pair of stochastic processes---i.e., one for each prospect---and the subsequent comparison of their final values is repeated until the sum of round wins for one of the two prospects reaches a threshold $\theta \in \mathbb{N}$.
Similar to the switching probability, it is assumed that the threshold is adopted prior the start of the information-processing chain and is fixed throughout.
For $P'^X$, then, we obtain a sequence of independent comparisons $Z^X_1, Z^X_2, ...$ and model the respective evidence accumulation process as a random walk $\{D^X_m\}_{m \in \mathbb{N}}$ defined by 
$$
D^X_m = \sum_{i = 1}^m Z^X_i, \quad m \in \mathbb{N}
\;,
\tag{9}
$$
resembling a dynamic decision variable.
Note that since (8) describes a Bernoulli trial, the random walk in (9) is based on a Bernoulli process.

Finally, note that the described model assumes two random walks---i.e., one for each prospect---approaching the same threshold, which is the absolute number of mean comparisons a prospect must win in order to be chosen.
However, one may also define a single random walk, resembling a common dynamic decision variable, which approaches a positive threshold $\theta^+$ and a negative threshold $\theta^-$.
Therefore, (8) may be adapted such that $Z$ takes the value $-1$ instead of $0$, such that the threshold determines how many more mean comparisons a prospect must win in order to be chosen.

# Model Predictions

Below, the key predictions concerning the effects of the model parameters $\psi$ and $\theta$ on the length of the sampling and accumulation process and the resulting choice are outlined.
Moreover, previous research has shown that differences as to how information about the properties of prospects is processed can translate to characteristic shapes of cumulative prospect theory's (CPT) value and weighting function [see @pachurHowTwainCan2017; @zilkerNonlinearProbabilityWeighting2021].
Respectively, this section also contains predictions about the relations between $\psi$ and $\theta$ and the CPT parameters.
A summary of the model predictions can be found in Table\ \@ref(tab:predictions).

## Length of Information-Processing Chains

The required number of independent mean comparisons increases with thresholds $\theta$.
Since each comparison is based on its own pair of stochastic processes, an increase in $\theta$ translates to a larger number of required stochastic processes and, ceteris paribus, therefore also to a larger overall number of outcomes sampled over the course of a sampling and accumulation process.
In turn, each stochastic process is predicted to terminate faster, the higher the probability that the subsequent outcome is sampled from the other prospect, determined by the switching probability $\psi$.
Accordingly, an increase in $\psi$ leads to shorter stochastic processes, which translate, ceteris paribus, to a smaller overall number of sampled outcomes.

## Final Choice

Because the switching probability $\psi$ is predicted to alter the length of the stochastic processes underlying each mean comparison, $\psi$ should affect whether low-probability outcomes contribute to the majority of the required mean comparisons about as much as would be warranted based on their objective probabilities, or whether they contribute rather less.
That is, for short stochastic processes, the binomial distribution associated with small-probability outcomes is positive skewed, therefore causing the relative frequencies with which these outcomes occur over the course of such a process to be smaller rather than larger than their latent objective probabilities.
Respectively, if the small-probability outcome is smaller (larger) than the EV, high values of $\psi$---and therefore pairs of short stochastic processes---cause the mean to be an inflated (deflated) estimate of the EV.
Such an inflation and deflation of the means can then translate to an as-if underweighting of rare outcomes.
Specifically, if the mean of the prospect with the lower (higher) EV is inflated (deflated), the outcome of the comparison between the means may be a reversal---in sign---of the outcome of a comparison between the EVs, leading to choices that take the form of an as-if underweighting of rare outcomes.
Yet, if the mean of the prospect with the lower (higher) EV is deflated (inflated), the difference between the means should be larger than the differences between the EV, which can simplify EV maximization [see @hertwigDecisionsExperienceWhy2010].
In turn, for low values of $\psi$---and therefore pairs of long stochastic processes---the means should closely correspond to the EVs, leading to choices that are in line with EV maximization.
To summarize the choice predictions for $\psi$ and simplify: 
If the prospect that is more likely to return a higher outcome most of the time does not also possess the higher EV, high values of $\psi$ are predicted to cause an as-if underweighting of rare outcomes.
Respectively, while these predictions emphasize the close link between small samples and the underweighting of rare outcomes, they suggest that the reliance on small samples can serve the accumulation of evidence for differences in a latent property.

More so, for an increasing sequence of independent comparisons between single outcomes sampled from two prospects, the relative winning frequency of a prospect should converge against the objective probability of producing a higher outcome in each comparison.
Accordingly, the as-if underweighting of rare outcomes should become more robust as the thresholds $\theta$ increase.

## Cumulative Prospect Theory

Before elaborating how the above model predictions may translate into estimates of the CPT parameters, the following section reviews how CPT in general and its weighting function in particular capture deviations from EV maximization.

### Formal Review of CPT
